---
title: "stat418hw4"
author: "Jixuan Li"
date: "6/7/2017"
output: html_document
---

```{r}
if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }
if (! ("methods" %in% rownames(installed.packages()))) { install.packages("methods") }
if (! ("statmod" %in% rownames(installed.packages()))) { install.packages("statmod") }
if (! ("stats" %in% rownames(installed.packages()))) { install.packages("stats") }
if (! ("graphics" %in% rownames(installed.packages()))) { install.packages("graphics") }
if (! ("RCurl" %in% rownames(installed.packages()))) { install.packages("RCurl") }
if (! ("jsonlite" %in% rownames(installed.packages()))) { install.packages("jsonlite") }
if (! ("tools" %in% rownames(installed.packages()))) { install.packages("tools") }
if (! ("utils" %in% rownames(installed.packages()))) { install.packages("utils") }
install.packages("h2o", type="source", repos=(c("http://h2o-release.s3.amazonaws.com/h2o/rel-ueno/8/R")))
library(h2o)
localH2O = h2o.init(nthreads=-1)
demo(h2o.kmeans)
```

```{r}
dx <- h2o.importFile("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep=",", header=F, col.names=c('age','workclass','fnlwgt','education','EducationNo','MaritalStatus','occupation','relationship','race','sex','CapitalGain','CapitalLoss','HoursPerWeek','NativeCountry','salary'))
```

Here, we split the dataset into training set, testing set and validation set. It produced 0.7/0.2/0.1 for the original adult dataset to initiate train/validation/test dataset.
```{r}
dx_split <- h2o.splitFrame(dx, ratios = c(0.7, 0.2), seed = 123)
dx_train <- dx_split[[1]]
dx_valid <- dx_split[[2]]
dx_test <- dx_split[[3]]
Xnames <- names(dx_train)[which(names(dx_train)!="salary")]
```

Model from hw3: logistic regression

```{r}
library(h2o)

h2o.init(nthreads=-1)

dx <- h2o.importFile("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep=",", header=F, col.names=c('age','workclass','fnlwgt','education','EducationNo','MaritalStatus','occupation','relationship','race','sex','CapitalGain','CapitalLoss','HoursPerWeek','NativeCountry','salary'))

dx_split <- h2o.splitFrame(dx, ratios = 0.7, seed = 123)
dx_train <- dx_split[[1]]
dx_test <- dx_split[[2]]


Xnames <- names(dx_train)[which(names(dx_train)!="salary")]

system.time({
  md <- h2o.glm(x = Xnames, y = "salary", training_frame = dx_train, 
                family = "binomial", alpha = 1, lambda = 0)
})



h2o.auc(h2o.performance(md, dx_test))


md
```

Model from hw3: Random Forest

```{r}
library(h2o)

h2o.init(nthreads=-1)

dx <- h2o.importFile("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep=",", header=F, col.names=c('age','workclass','fnlwgt','education','EducationNo','MaritalStatus','occupation','relationship','race','sex','CapitalGain','CapitalLoss','HoursPerWeek','NativeCountry','salary'))

dx_split <- h2o.splitFrame(dx, ratios = 0.6, seed = 123)
dx_train <- dx_split[[1]]
dx_test <- dx_split[[2]]


Xnames <- names(dx_train)[which(names(dx_train)!="salary")]

system.time({
  md <- h2o.randomForest(x = Xnames, y = "salary", training_frame = dx_train, ntrees = 500)
})



h2o.auc(h2o.performance(md, dx_test))


md
```

Model: neural network
Now, we perform the neural network algorithm on the training set.
```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", stopping_tolerance = 0) 
})
h2o.performance(md, dx_test)@metrics$AUC
```

```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(50,50,50,50), 
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", stopping_tolerance = 0) 
})
h2o.performance(md, dx_test)@metrics$AUC
```

We keep other conditions all the same, and set the adaptive learning rate to FALSE,
```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(50,50,50,50), 
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", adaptive_rate = FALSE, stopping_tolerance = 0)
})
h2o.performance(md, dx_test)@metrics$AUC
```

Adding to the last condition, we set the initial momentum to 0.2 at the beginning of the training to 0.2. The number of training samples for which momentum increase to 10000, and the set the final momentum which the ramp is over to 0.99,
```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(50,50,50,50), 
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", adaptive_rate = FALSE, momentum_start = 0.2, momentum_ramp = 10000, momentum_stable = 0.99, stopping_tolerance = 0)
})
h2o.performance(md, dx_test)@metrics$AUC
```

We try the L1 and L2 regulazations to different values.
```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(200,200), l1 = 1e-5, l2 = 1e-5, 
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", stopping_tolerance = 0) 
})
h2o.performance(md, dx_test)@metrics$AUC
```

For the following two tests, we change the different adaptive learning rate decay (rho) and smoothing (epsilon) factors to different values,
```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(200,200), l1 = 1e-5, l2 = 1e-5, 
            rho = 0.95, epsilon = 1e-06,
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", stopping_tolerance = 0) 
})
h2o.performance(md, dx_test)@metrics$AUC
```

```{r}
system.time({
  md <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, validation_frame = dx_valid,
            activation = "Rectifier", hidden = c(200,200), l1 = 1e-5, l2 = 1e-5, 
            rho = 0.9999, epsilon = 1e-09,
            epochs = 100, stopping_rounds = 2, stopping_metric = "AUC", stopping_tolerance = 0) 
})
h2o.performance(md, dx_test)@metrics$AUC
```

Model 2: 

```{r}
library(h2o)

h2o.init(nthreads=-1)

dx <- h2o.importFile("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep=",", header=F, col.names=c('age','workclass','fnlwgt','education','EducationNo','MaritalStatus','occupation','relationship','race','sex','CapitalGain','CapitalLoss','HoursPerWeek','NativeCountry','salary'))

dx_split <- h2o.splitFrame(dx, ratios = c(0.6,0.2), seed = 123)
dx_train <- dx_split[[1]]
dx_valid <- dx_split[[2]]
dx_test <- dx_split[[3]]


Xnames <- names(dx_train)[which(names(dx_train)!="salary")]


hyper_params <- list( ntrees = 10000,  ## early stopping
                     max_depth = 5:15, 
                     min_rows = c(1,3,10,30,100),
                     learn_rate = c(0.01,0.03,0.1),  
                     learn_rate_annealing = c(0.99,0.995,1,1),
                     sample_rate = c(0.4,0.7,1,1),
                     col_sample_rate = c(0.7,1,1),
                     nbins = c(30,100,300),
                     nbins_cats = c(64,256,1024)
)

search_criteria <- list( strategy = "RandomDiscrete",
                        max_runtime_secs = 10*3600,
                        max_models = 100
)

system.time({
  mds <- h2o.grid(algorithm = "gbm", grid_id = "grd",
                  x = Xnames, y = "dep_delayed_15min", training_frame = dx_train,
                  validation_frame = dx_valid,
                  hyper_params = hyper_params,
                  search_criteria = search_criteria,
                  stopping_metric = "AUC", stopping_tolerance = 1e-3, stopping_rounds = 2,
                  seed = 123)
})



mds_sort <- h2o.getGrid(grid_id = "grd", sort_by = "auc", decreasing = TRUE)
mds_sort

md_best <- h2o.getModel(mds_sort@model_ids[[1]])
summary(md_best)

h2o.auc(h2o.performance(md_best, dx_test))
```

Model 3: esembles

```{r}
library(h2o)

h2o.init(nthreads=-1)

dx <- h2o.importFile("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data", sep=",", header=F, col.names=c('age','workclass','fnlwgt','education','EducationNo','MaritalStatus','occupation','relationship','race','sex','CapitalGain','CapitalLoss','HoursPerWeek','NativeCountry','salary'))

dx_split <- h2o.splitFrame(dx, ratios = 0.7, seed = 123)
dx_train <- dx_split[[1]]
dx_test <- dx_split[[2]]


Xnames <- setdiff(names(dx_train),"salary")



system.time({
  md1 <- h2o.glm(x = Xnames, y = "salary", training_frame = dx_train, 
                family = "binomial", 
                alpha = 1, lambda = 0,
                seed = 123,
                nfolds = 5, fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE)
})

system.time({
  md2 <- h2o.randomForest(x = Xnames, y = "salary", training_frame = dx_train, 
                ntrees = 300,
                seed = 123,
                nfolds = 5, fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE)
})


system.time({
  md3 <- h2o.gbm(x = Xnames, y = "salary", training_frame = dx_train, distribution = "bernoulli", 
                ntrees = 200, max_depth = 10, learn_rate = 0.1, 
                nbins = 100, seed = 123,
                nfolds = 5, fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE)    
})

system.time({
  md4 <- h2o.deeplearning(x = Xnames, y = "salary", training_frame = dx_train, 
            epochs = 5,
            seed = 123,
            nfolds = 5, fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE) 
})



md_ens <- h2o.stackedEnsemble(x = Xnames, y = "salary", training_frame = dx_train, 
                    base_models = list(md1@model_id, md2@model_id, md3@model_id, md4@model_id))


h2o.auc(h2o.performance(md1, dx_test))
h2o.auc(h2o.performance(md2, dx_test))
h2o.auc(h2o.performance(md3, dx_test))
h2o.auc(h2o.performance(md4, dx_test))
h2o.auc(h2o.performance(md_ens, dx_test))


h2o.getModel(md_ens@model$metalearner$name)@model$coefficients_table
```
